Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses

### 2022-09-18 RR_manuscript/Draft_RR.Rmd
include desc table (draft), fix typos

Speech is a complex signal that is characterized by structurally different acoustic landmarks distributed throughout different temporal domains.
Thus, choosing how to measure a communictive intention of interest is an important analytic step.
Take for example the following sentence in (1).

@@

```{r}
#| label: team-descriptives-table
#| echo: false
#| eval: false
team_desc_df %>% 
  mutate(Mean = glue::glue("{mean_size} ({sd_size})")) %>% 
  select(Stage = stage, n = n_teams, Mean, Median = median_size) %>% 
  knitr::kable(align = c("l", "r", "r", "r"), 
               caption = "Caption test", 
               label = "team-descriptives-table")
```

`r team_desc_list$Initial$n_teams$val %>% xfun::numbers_to_words() %>% stringr::str_to_title()` teams initially signed up to participate in the study.
`r (team_desc_list$Initial$n_teams$val - team_desc_list$Final$n_teams$val) %>% xfun::numbers_to_words() %>% stringr::str_to_title()` of the signed-up teams dropped out during the analysis phase.
One Team Coordinator, Wim van Dommelen (who agreed to their feedback being shared), gave the following reason: "The reason is that [...] the project's approach is too much top-down instead of bottom-up. [...] I would rather prefer to explore the collected utterances and run different types of acoustic analyses, which would then be followed by statistical evaluation of the data."
We find this contribution to clearly speak to the main objective of the current study: investigate researcher's degrees of freedom in the speech sciences.
Based on our personal experience with research in the field, it is common practice to test as many different types of models, using as many different types of measurements, to asnwer one research question.
The remaining `r team_desc_list$Final$n_teams$val` teams submitted their analysis by the established deadline.
These are the teams that are included in this study.

@@

`r round(window_n[[1,1]],2) * 100`\% of models obtained acoustic measures at the level of the segment (e.g. comparing the acoustic profile of a vowel), `r round(window_n[[2,1]],2) * 100`\% from the word level (e.g. comparing the acoustic profile of *Banane* 'banana'), `r round(window_n[[3,1]],2) * 100`\% at the level of the phrase (e.g. the noun phrase including determiner and adjective, e.g. "the green banana"), `r round(window_n[[4,1]],2) * 100`\% from the whole sentence, and `r round(window_n[[5,1]],2) * 100`\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of `r nrow(unique_acou_oper)` different measurement specifications.
For example, if we consider those analyses who targetfundamental frequency (f0), we find that it is operationalized in a myriad of ways including the minimum, the maximum, the mean, the median, as a range in an interval or a ratio between two intervals. 
The measurement is sometimes taken from the interval of a vowel in the article, the adjective or noun; it is sometimes taken from the word interval of article, adjective or noun; or it is taken from either the noun phrase interval or the entire sentence. 
Some of these measures were normalized relative to other elements in the sentence or relative to the speaker.

@@

## Caveats
Our study has several limitations that need to be considered when evaluating our results. 
Importantly, there is potential for bias. 
The coordinating authors have engaged with metascientific research before and have been actively involved in methodological debates about scientific practices including transparency and statistical methods. 
We have in the past used the lack of standardized analytical approaches as an argument for proposing behavior and policy changes in the field. 
This might have biased our own judgment during the analysis which itself came with many researcher degrees of freedom. 
We hope we were able to make these degrees of freedom as well as the timing and reasoning of our analytical choices at least detectable. 
However, our results should be independently re-analysed and replicated using a different research question.
Moreover, our sample is an opportunity sample. 
We have advertised the project through online platforms which might have led to the exclusion of certain potential researcher groups.
While the number of participating teams is larger than most earlier crowdsourcing projects, it is likely to be too small to estimate for example the impact of certain predictors reliably.
Since predictor levels were not systematically distributed across teams, our estimates are characterized by large uncertainty. 
This uncertainty is possibly further inflated by the fact that he research question presented to the teams was vague: *Do speakers acoustically modify utterances to signal atypical word combinations?* 
Interpreting the research question/hypothesis differently in terms of its statistical consequences has recently been shown to explain some variation between analysis teams in many-analysis projects [@auspurg2021has]. 
However, we consider this very underspecification of research hypotheses in the field of speech science (and beyond, see @scheel2022most) a common phenomenon. 
For example, researchers seem to have not yet agreed on how to acoustically measure cross linguistically common phenomena such as word stress [e.g. @gordon2017acoustic]. 
In light of these possible sources of bias, a field can benefit from explicit positionality statement [e.g., @fox2021open]
Researchers do not analyse data in a vaccuum.
It is important recognize and disclose one’s positionality, i.e., a reflection about how educational background, social identity, power, experience and context might influence researchers approaches and interpretations.
Despite these all of these possible sources of bias and uncertainty, our study suggest that at a minimum we can expect the between-team and between-analysis variability that we have observed, which is quite substantial and possibly amplifying biases. 
Beyond the potential for bias, our study lacks generalizability. 
The generalizability of our findings to other disciplines on the one hand and specifically other sub-disciplines of the language sciences is, of course, limited.
We focused on quantitative analyses that require operationalization of a multidimensional signal. 
While some of our results can be informative for other disciplines that work with speech or video signals, they might not be very informative for expected variation between analysts working in qualitative research fields.
This is an important point to make because the cognitive sciences in general, and the language sciences in particular have many research areas that are explored with qualitative methods rather than quantitatively. 
It is conceivable that the discussed issues apply differently or not at all to qualitative data analysis.

>>>>

Speech is a complex signal that is characterized by structurally different acoustic landmarks distributed throughout different temporal domains.
Thus, choosing how to measure a communicative intention of interest is an important analytic step.
Take for example the following sentence in (1).

@@

\phantom{.}

```{r}
#| label: team-descriptives-table
#| echo: false
#| results: 'asis'
#| eval: true

# team_desc_df %>% 
#   mutate(Mean = glue::glue("{mean_size} ({sd_size})")) %>% 
#   select(Stage = stage, n = n_teams, Mean, Median = median_size) %>% 
#   knitr::kable(align = c("l", "r", "r", "r"), 
#                caption = "Caption test", 
#                label = "team-descriptives-table")

# Team descriptives
tbl_team_desc <- msa_models %>% 
  select(animal, years_from_phd, prior_belief, phon_rating, stat_rating, 
         all_rating) %>% 
  group_by(animal) %>% 
  summarize(across(years_from_phd:all_rating, mean)) %>% 
  pivot_longer(cols = -animal, names_to = "metric", values_to = "vals") %>% 
  na.omit() %>% 
  group_by(metric) %>% 
  summarize(min = min(vals), max = max(vals), avg = mean(vals), sd = sd(vals)) %>% 
  mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(desc = "team_desc", 
    metric = fct_relevel(metric, "years_from_phd"), 
   `Mean (SD)` = glue::glue("{avg} ({sd})"), 
    Metric = case_when(
      metric == "years_from_phd" ~ "Years since PhD", 
      metric == "stat_rating" ~ "Statistical ability", 
      metric == "prior_belief" ~ "Prior belief of $H_{1}$", 
      metric == "phon_rating" ~ "Phon rating", 
      metric == "all_rating" ~ "all rating"), 
    label = glue::glue("{desc}_{metric}")) %>% 
  select(label, desc, Metric, Min. = min, Max = max, `Mean (SD)`, avg, sd) %>% 
  add_row(label = " ", desc = "team_desc", Metric = " ", 
          Min. = NA, Max = NA, `Mean (SD)` = NA, avg = NA, sd = NA) %>% 
  mutate(across(-c("label", "desc", "Metric"), ~replace(., is.na(.), " ")))

# Model descriptives 
tbl_model_desc <- msa_models %>% 
  select(animal, n_predictors:n_random, n_models) %>% 
  group_by(animal) %>% 
  summarize(across(everything(), mean)) %>% 
  pivot_longer(cols = -animal, names_to = "metric", values_to = "vals") %>% 
  na.omit() %>% 
  group_by(metric) %>% 
  summarize(min = min(vals), max = max(vals), avg = mean(vals), sd = sd(vals)) %>% 
  mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(desc = "model_desc", 
    metric = fct_relevel(metric, "n_models"), 
    `Mean (SD)` = glue::glue("{avg} ({sd})"), 
    Metric = case_when(
      metric == "n_coefficients" ~ "n. coefficients", 
      metric == "n_models" ~ "n. models", 
      metric == "n_predictors" ~ "n. predictors", 
      metric == "n_random" ~ "n. random effects", 
      metric == "n_random_int" ~ "n. random intercepts", 
      metric == "n_random_slo" ~ "n. random slopes", 
      metric == "n_random_smooth" ~ "n. random smooths",
      metric == "n_terms" ~ "n. terms"), 
    label = glue::glue("{desc}_{metric}")) %>% 
  select(label, desc, Metric, Min. = min, Max = max, `Mean (SD)`, avg, sd) %>% 
  mutate(across(-c("desc", "Metric"), as.character))

# Combine in single tibble (for printing in prose and table)
tbl_team_model_desc <- bind_rows(tbl_team_desc, tbl_model_desc) %>% 
  mutate(desc = case_when(
    Metric != "all rating" & desc == "team_desc" ~ " ", 
    Metric != "n. coefficients" & desc == "model_desc" ~ " ", 
    TRUE ~ desc), 
    desc = case_when(
      desc == "team_desc" ~ "Teams", 
      desc == "model_desc" ~ "Models", 
      TRUE ~ " "))

tbl_team_model_desc %>% 
  select(Descriptives = desc, Metric, Min., Max, `Mean (SD)`) %>% 
  knitr::kable(format = "latex", booktabs = T, escape = F, 
    linesep = "", align = c("l", "l", "r", "r", "r"))

```

`r team_desc_list$Initial$n_teams$val %>% xfun::numbers_to_words() %>% stringr::str_to_sentence()` teams initially signed up to participate in the study.
`r (team_desc_list$Initial$n_teams$val - team_desc_list$Final$n_teams$val) %>% xfun::numbers_to_words() %>% stringr::str_to_sentence()` of the signed-up teams dropped out during the analysis phase.
One Team Coordinator, Wim van Dommelen (who agreed to their feedback being shared), gave the following reason: "The reason is that [...] the project's approach is too much top-down instead of bottom-up. [...] I would rather prefer to explore the collected utterances and run different types of acoustic analyses, which would then be followed by statistical evaluation of the data."
We find this contribution to clearly speak to the main objective of the current study: investigate researcher's degrees of freedom in the speech sciences.
Based on our personal experience with research in the field, it is common practice to test as many different types of models, using as many different types of measurements, to answer one research question.
The remaining `r team_desc_list$Final$n_teams$val` teams submitted their analysis by the established deadline.
These are the teams that are included in this study.

@@

`r (round(window_n[[1,1]],2) * 100) %>% xfun::numbers_to_words() %>% stringr::str_to_sentence()` percent of models obtained acoustic measures at the level of the segment (e.g. comparing the acoustic profile of a vowel), `r round(window_n[[2,1]],2) * 100`\% from the word level (e.g. comparing the acoustic profile of *Banane* 'banana'), `r round(window_n[[3,1]],2) * 100`\% at the level of the phrase (e.g. the noun phrase including determiner and adjective, e.g. "the green banana"), `r round(window_n[[4,1]],2) * 100`\% from the whole sentence, and `r round(window_n[[5,1]],2) * 100`\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of `r nrow(unique_acou_oper)` different measurement specifications.
For example, if we consider those analyses that target fundamental frequency (f0), we find that it is operationalized in a myriad of ways including the minimum, the maximum, the mean, the median, as a range in an interval or a ratio between two intervals. 
The measurement is sometimes taken from the interval of a vowel in the article, the adjective or noun; it is sometimes taken from the word interval of article, adjective or noun; or it is taken from either the noun phrase interval or the entire sentence. 
Some of these measures were normalized relative to other elements in the sentence or relative to the speaker.

@@

## Caveats
Our study has several limitations that need to be considered when evaluating our results. 
Importantly, there is potential for bias. 
The coordinating authors have engaged with metascientific research before and have been actively involved in methodological debates about scientific practices including transparency and statistical methods. 
We have in the past used the lack of standardized analytical approaches as an argument for proposing behavior and policy changes in the field. 
This might have biased our own judgment during the analysis which itself came with many researcher degrees of freedom. 
We hope we were able to make these degrees of freedom as well as the timing and reasoning of our analytical choices at least detectable. 
However, our results should be independently re-analysed and replicated using a different research question.
Moreover, our sample is an opportunity sample. 
We have advertised the project through online platforms which might have led to the exclusion of certain potential researcher groups.
While the number of participating teams is larger than most earlier crowdsourcing projects, it is likely to be too small to estimate for example the impact of certain predictors reliably.
Since predictor levels were not systematically distributed across teams, our estimates are characterized by large uncertainty. 
This uncertainty is possibly further inflated by the fact that he research question presented to the teams was vague: *Do speakers acoustically modify utterances to signal atypical word combinations?* 
Interpreting the research question/hypothesis differently in terms of its statistical consequences has recently been shown to explain some variation between analysis teams in many-analysis projects [@auspurg2021has]. 
However, we consider this very underspecification of research hypotheses in the field of speech science (and beyond, see @scheel2022most) a common phenomenon. 
For example, researchers seem to have not yet agreed on how to acoustically measure cross linguistically common phenomena such as word stress [e.g. @gordon2017acoustic]. 
In light of these possible sources of bias, a field can benefit from explicit positionality statements [e.g., @fox2021open]
Researchers do not analyse data in a vaccuum.
It is important recognize and disclose one’s positionality, i.e., a reflection about how educational background, social identity, power, experience and context might influence researchers approaches and interpretations.
Despite these all of these possible sources of bias and uncertainty, our study suggest that at a minimum we can expect the between-team and between-analysis variability that we have observed, which is quite substantial and possibly amplifying biases. 
Beyond the potential for bias, our study lacks generalizability. 
The generalizability of our findings to other disciplines on the one hand and specifically other sub-disciplines of the language sciences is, of course, limited.
We focused on quantitative analyses that require operationalization of a multidimensional signal. 
While some of our results can be informative for other disciplines that work with speech or video signals, they might not be very informative for expected variation between analysts working in qualitative research fields.
This is an important point to make because the cognitive sciences in general, and the language sciences in particular have many research areas that are explored with qualitative methods rather than quantitatively. 
It is conceivable that the discussed issues apply differently or not at all to qualitative data analysis.

### 2022-09-18 RR_manuscript/Draft_RR.tex
include desc table (draft), fix typos

Eighty-Four teams initially signed up to participate in the study.
Fifty-Four of the signed-up teams dropped out during the analysis phase.
One Team Coordinator, Wim van Dommelen (who agreed to their feedback being shared), gave the following reason: ``The reason is that {[}\ldots{]} the project's approach is too much top-down instead of bottom-up. {[}\ldots{]} I would rather prefer to explore the collected utterances and run different types of acoustic analyses, which would then be followed by statistical evaluation of the data.''
We find this contribution to clearly speak to the main objective of the current study: investigate researcher's degrees of freedom in the speech sciences.
Based on our personal experience with research in the field, it is common practice to test as many different types of models, using as many different types of measurements, to asnwer one research question.
The remaining 30 teams submitted their analysis by the established deadline.
These are the teams that are included in this study.

@@

45\% of models obtained acoustic measures at the level of the segment (e.g.~comparing the acoustic profile of a vowel), 43\% from the word level (e.g.~comparing the acoustic profile of \emph{Banane} `banana'), 3\% at the level of the phrase (e.g.~the noun phrase including determiner and adjective, e.g.~``the green banana''), 4\% from the whole sentence, and 4\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of 52 different measurement specifications.
For example, if we consider those analyses who targetfundamental frequency (f0), we find that it is operationalized in a myriad of ways including the minimum, the maximum, the mean, the median, as a range in an interval or a ratio between two intervals.
The measurement is sometimes taken from the interval of a vowel in the article, the adjective or noun; it is sometimes taken from the word interval of article, adjective or noun; or it is taken from either the noun phrase interval or the entire sentence.
Some of these measures were normalized relative to other elements in the sentence or relative to the speaker.

@@

Our study has several limitations that need to be considered when evaluating our results.
Importantly, there is potential for bias.
The coordinating authors have engaged with metascientific research before and have been actively involved in methodological debates about scientific practices including transparency and statistical methods.
We have in the past used the lack of standardized analytical approaches as an argument for proposing behavior and policy changes in the field.
This might have biased our own judgment during the analysis which itself came with many researcher degrees of freedom.
We hope we were able to make these degrees of freedom as well as the timing and reasoning of our analytical choices at least detectable.
However, our results should be independently re-analysed and replicated using a different research question.
Moreover, our sample is an opportunity sample.
We have advertised the project through online platforms which might have led to the exclusion of certain potential researcher groups.
While the number of participating teams is larger than most earlier crowdsourcing projects, it is likely to be too small to estimate for example the impact of certain predictors reliably.
Since predictor levels were not systematically distributed across teams, our estimates are characterized by large uncertainty.
This uncertainty is possibly further inflated by the fact that he research question presented to the teams was vague: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}
Interpreting the research question/hypothesis differently in terms of its statistical consequences has recently been shown to explain some variation between analysis teams in many-analysis projects \citep{auspurg2021has}.
However, we consider this very underspecification of research hypotheses in the field of speech science (and beyond, see \citet{scheel2022most}) a common phenomenon.
For example, researchers seem to have not yet agreed on how to acoustically measure cross linguistically common phenomena such as word stress \citep[e.g.][]{gordon2017acoustic}.
In light of these possible sources of bias, a field can benefit from explicit positionality statement \citep[e.g.,][]{fox2021open}
Researchers do not analyse data in a vaccuum.
It is important recognize and disclose one's positionality, i.e., a reflection about how educational background, social identity, power, experience and context might influence researchers approaches and interpretations.
Despite these all of these possible sources of bias and uncertainty, our study suggest that at a minimum we can expect the between-team and between-analysis variability that we have observed, which is quite substantial and possibly amplifying biases.
Beyond the potential for bias, our study lacks generalizability.
The generalizability of our findings to other disciplines on the one hand and specifically other sub-disciplines of the language sciences is, of course, limited.
We focused on quantitative analyses that require operationalization of a multidimensional signal.
While some of our results can be informative for other disciplines that work with speech or video signals, they might not be very informative for expected variation between analysts working in qualitative research fields.
This is an important point to make because the cognitive sciences in general, and the language sciences in particular have many research areas that are explored with qualitative methods rather than quantitatively.
It is conceivable that the discussed issues apply differently or not at all to qualitative data analysis.

>>>>

\phantom{.}

\begin{tabular}{llrrr}
\toprule
Descriptives & Metric & Min. & Max & Mean (SD)\\
\midrule
Teams & all rating & 39 & 88.67 & 69.95 (14.34)\\
 & Phon rating & 41.25 & 88.33 & 71.51 (13.92)\\
 & Prior belief of $H_{1}$ & 48.5 & 92 & 69.19 (11.33)\\
 & Statistical ability & 33 & 93.33 & 69.46 (16.52)\\
 & Years since PhD & -3.75 & 12.38 & 4.77 (4.16)\\
 &  &  &  & \\
Models & n. coefficients & 1 & 11 & 3.5 (2.75)\\
 & n. models & 1 & 16 & 3.67 (3.44)\\
 & n. predictors & 1 & 4.2 & 1.59 (0.85)\\
 & n. random effects & 0 & 10 & 2.47 (2.05)\\
 & n. random intercepts & 0 & 10 & 1.91 (1.77)\\
 & n. random slopes & 0 & 4 & 0.63 (1.07)\\
 & n. random smooths & 0 & 0 & 0 (0)\\
 & n. terms & 1 & 4.27 & 1.78 (1.08)\\
\bottomrule
\end{tabular}

Eighty-four teams initially signed up to participate in the study.
Fifty-four of the signed-up teams dropped out during the analysis phase.
One Team Coordinator, Wim van Dommelen (who agreed to their feedback being shared), gave the following reason: ``The reason is that {[}\ldots{]} the project's approach is too much top-down instead of bottom-up. {[}\ldots{]} I would rather prefer to explore the collected utterances and run different types of acoustic analyses, which would then be followed by statistical evaluation of the data.''
We find this contribution to clearly speak to the main objective of the current study: investigate researcher's degrees of freedom in the speech sciences.
Based on our personal experience with research in the field, it is common practice to test as many different types of models, using as many different types of measurements, to answer one research question.
The remaining 30 teams submitted their analysis by the established deadline.
These are the teams that are included in this study.

@@

Forty-five percent of models obtained acoustic measures at the level of the segment (e.g.~comparing the acoustic profile of a vowel), 43\% from the word level (e.g.~comparing the acoustic profile of \emph{Banane} `banana'), 3\% at the level of the phrase (e.g.~the noun phrase including determiner and adjective, e.g.~``the green banana''), 4\% from the whole sentence, and 4\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of 52 different measurement specifications.
For example, if we consider those analyses that target fundamental frequency (f0), we find that it is operationalized in a myriad of ways including the minimum, the maximum, the mean, the median, as a range in an interval or a ratio between two intervals.
The measurement is sometimes taken from the interval of a vowel in the article, the adjective or noun; it is sometimes taken from the word interval of article, adjective or noun; or it is taken from either the noun phrase interval or the entire sentence.
Some of these measures were normalized relative to other elements in the sentence or relative to the speaker.

@@

Our study has several limitations that need to be considered when evaluating our results.
Importantly, there is potential for bias.
The coordinating authors have engaged with metascientific research before and have been actively involved in methodological debates about scientific practices including transparency and statistical methods.
We have in the past used the lack of standardized analytical approaches as an argument for proposing behavior and policy changes in the field.
This might have biased our own judgment during the analysis which itself came with many researcher degrees of freedom.
We hope we were able to make these degrees of freedom as well as the timing and reasoning of our analytical choices at least detectable.
However, our results should be independently re-analysed and replicated using a different research question.
Moreover, our sample is an opportunity sample.
We have advertised the project through online platforms which might have led to the exclusion of certain potential researcher groups.
While the number of participating teams is larger than most earlier crowdsourcing projects, it is likely to be too small to estimate for example the impact of certain predictors reliably.
Since predictor levels were not systematically distributed across teams, our estimates are characterized by large uncertainty.
This uncertainty is possibly further inflated by the fact that he research question presented to the teams was vague: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}
Interpreting the research question/hypothesis differently in terms of its statistical consequences has recently been shown to explain some variation between analysis teams in many-analysis projects \citep{auspurg2021has}.
However, we consider this very underspecification of research hypotheses in the field of speech science (and beyond, see \citet{scheel2022most}) a common phenomenon.
For example, researchers seem to have not yet agreed on how to acoustically measure cross linguistically common phenomena such as word stress \citep[e.g.][]{gordon2017acoustic}.
In light of these possible sources of bias, a field can benefit from explicit positionality statements \citep[e.g.,][]{fox2021open}
Researchers do not analyse data in a vaccuum.
It is important recognize and disclose one's positionality, i.e., a reflection about how educational background, social identity, power, experience and context might influence researchers approaches and interpretations.
Despite these all of these possible sources of bias and uncertainty, our study suggest that at a minimum we can expect the between-team and between-analysis variability that we have observed, which is quite substantial and possibly amplifying biases.
Beyond the potential for bias, our study lacks generalizability.
The generalizability of our findings to other disciplines on the one hand and specifically other sub-disciplines of the language sciences is, of course, limited.
We focused on quantitative analyses that require operationalization of a multidimensional signal.
While some of our results can be informative for other disciplines that work with speech or video signals, they might not be very informative for expected variation between analysts working in qualitative research fields.
This is an important point to make because the cognitive sciences in general, and the language sciences in particular have many research areas that are explored with qualitative methods rather than quantitatively.
It is conceivable that the discussed issues apply differently or not at all to qualitative data analysis.
